---
title: "MaxQuant_Analysis"
author: "Sarah Arcos"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(limma)
library(ggpubr)
library(VennDiagram)
library(clusterProfiler)
library(openxlsx)
```

# This markdown processes the peptides.txt output from MaxQuant to determine significantly enriched proteins from a VIR-CLASP experiment. 

## Step 0: Read in files and perform pre-processing
 - Select relevant columns
 - Remove reverse sequences and potential contaminants (but preserve benzonase (NUCA_SERMA))
 - Clean protein names
 - Set 0 values to NA
 - Split into one tibble for each condition tested
 
```{r, message = FALSE}
peptides_full <- read_tsv("MaxQuant/peptides.txt") %>%
  select(`Sequence`, Proteins, `Unique (Proteins)`, `Unique (Groups)`, `Leading razor protein`, contains("LFQ intensity")) %>%
  filter(`Unique (Proteins)` == "yes" &
           !str_detect(`Leading razor protein`, "REV") &
           !str_detect(`Leading razor protein`, "CON") |
           str_detect(`Proteins`, "SERMA")) %>%
  select(-c("Unique (Proteins)", "Unique (Groups)", "Leading razor protein")) %>%
  mutate(Proteins = case_when(
    Proteins %in% c("Capsid", "E1", "E2", "nsP1", "nsP2") ~ Proteins,
    TRUE ~ str_extract(Proteins, "[OPQ][0-9][A-Z0-9]{3}[0-9]|[A-NR-Z][0-9]([A-Z][A-Z0-9]{2}[0-9]){1,2}")
  ))

peptides_full[peptides_full == 0] <- NA
```

## Main: call this function on each condition

```{r, warning = FALSE}
main <- function(dataset, name) {
  
  results_dir <- paste(name, " Results", sep = "")
  dir.create(results_dir)
  
  data_1 <- step1(dataset)
  
  data_3 <- data_1 %>%
    step2() %>%
    step3()
  
  quant <- data_3 %>%
    step4_plot(name, results_dir)
  
  semi_quant <- data_1 %>%
    step5() %>%
    step6() %>%
    step7_plot(name, results_dir) %>%
    step8()
  
  final_table <- data_3 %>%
    step9(quant, semi_quant, results_dir)
  
  step10(data_1, final_table, name)
  
  final_table
}

```

## Step 1: Only consider proteins "with at least two quantified peptides" (CARIC paper)
 - After splitting into conditions above, some peptides might have all NAs, so need to remove

```{r}
step1 <- function(dataset){
  print("Executing Step 1")
  dataset %>%
    mutate(all_mean = rowMeans(.[,3:6], na.rm = T)) %>%
    filter(!is.na(all_mean)) %>%
    select(-all_mean) %>%
    group_by(Proteins) %>%
    add_tally(name = "Count") %>%
    filter(Count > 1) %>%
    select(-Count) %>%
    ungroup()
}

#step1_test <- step1(pep_hour0_plus)
```

## Step 2: Calculate peptide intensity ratios
 - Also clean up sample names and add some grouping columns
 - Calculate mean intensity from no4SU samples
 - Calculate peptide intensity ratios, then log2
 - Calculate log of peptide intensity ratios

```{r}
step2 <- function(dataset){
  print("Executing Step 2")
  dataset %>%
    mutate(ave_no4SU = rowMeans(.[,5:6], na.rm = T)) %>%
    select(1:4, 7) %>%
    gather(Sample, Intensity, -Sequence, -Proteins, -ave_no4SU) %>%
    mutate(Sample = str_remove(Sample, "LFQ intensity "),
           log_intensity_ratio = log2(Intensity/ave_no4SU)) %>%
    select(-ave_no4SU, -Intensity)
}

#step2_test <- step2(step1_test)
```

## Step 3: Calculate mean log ratio for each protein
 - Trimmed mean (20%): “The differential abundance of a protein group was estimated as the trimmed mean over all differential peptide abundance where 20% of the data are trimmed on both sides.”
 - From 2012 Castello
 
```{r}
step3 <- function(dataset){
  print("Executing Step 3")
  dataset %>%
    group_by(Proteins, Sample) %>%
    mutate(ave_log_ratio = mean(log_intensity_ratio, trim = 0.20, na.rm = TRUE)) %>%
    select(-log_intensity_ratio, -Sequence) %>%
    unique()
}

#step3_test <- step3(step2_test)
```

## Step 4: Limma moderated t test
 - Generate scatter plots
 - Significance cutoff is p.adjust < 0.01
 
```{r}
step4_plot <- function(dataset, name, results_dir) {
  print("Executing Step 4")
  limma_data <- dataset %>%
    ungroup() %>%
    spread(Sample, ave_log_ratio) %>%
    column_to_rownames(var = "Proteins") %>%
    lmFit() %>%
    eBayes() %>%
    topTable(number = Inf) %>%
    rownames_to_column(var = "Proteins")
  
  plot_data <- dataset %>%
    ungroup() %>%
    spread(Sample, ave_log_ratio) %>%
    left_join(limma_data, by = "Proteins") %>%
    mutate(Significant = case_when(
      adj.P.Val < 0.01  & AveExpr > (log2(5)) ~ "Yes",
      TRUE ~ "No"
    ))
  
  figure <- ggplot(plot_data, aes(x = pull(plot_data, var = 2), y = pull(plot_data, var = 3), color = Significant)) +
    geom_point(alpha = 0.3) +
    geom_hline(yintercept = 0) +
    geom_hline(yintercept = 1, color = "darkgreen", linetype = "dashed", size = 0.5) +
    geom_vline(xintercept = 0) +
    geom_vline(xintercept = 1, color = "darkgreen", linetype = "dashed", size = 0.5) +
    scale_color_manual(values = c("black", "red")) +
    theme_minimal() +
    labs(x = "Replicate 1", y = "Replicate 2", title = name)
  
  print(figure)
  ggsave(paste(results_dir, "/Quant_Scatters.pdf", sep = ""))
  
  limma_data
}

#step4_test <- step4_plot(step3_test, "Name Test", results_dir = "Figures")

```
 
## Step 5: Count peptide occurrences

```{r}
step5 <- function(dataset){
  print("Executing Step 5")
  temp <- dataset %>%
    gather(Sample, Intensity, -Proteins, -Sequence) %>%
    filter(!is.na(Intensity)) %>%
    mutate(Sample = str_remove(Sample, "LFQ intensity "),
          Condition = str_sub(Sample, start = 1, end = 5)
          ) %>%
    group_by(Sequence, Condition) %>%
    add_tally() %>%
    select(-Sample, -Intensity) %>%
    unique() %>%
    spread(Condition, n)
  
  temp[is.na(temp)] <- 0
  
  temp
}

#step5_test <- step5(step1_test)
```

## Step 6: Assign significance for semi-quant
 - Don't need to do all matrix positions, just the likely ones (bottom left 2x2 square)
 - Significance is for __peptides__ not the proteins. Protein "hit" determination comes later

```{r}
step6 <- function(dataset){
  print("Executing Step 6")
  fdrs <- dataset %>%
    ungroup() %>%
    mutate(mat_pos = str_c(.[[3]], .[[4]], sep = ":")) %>%
    group_by(mat_pos) %>%
    tally() %>%
    add_missing() %>%
    spread(mat_pos, n) %>%
    gather(mat_pos, count, `3:0`, `3:1`, `2:0`, `2:1`) %>%
    mutate(FDR = case_when(
      mat_pos == "3:0" ~ `0:2`/count,
      mat_pos == "3:1" ~ `1:2`/count,
      mat_pos == "2:0" ~ `0:2`/count,
      mat_pos == "2:1" ~ `1:2`/count
    )) %>%
    select(mat_pos, FDR)
  
  dataset %>%
    ungroup() %>%
    mutate(mat_pos = str_c(.[[3]], .[[4]], sep = ":")) %>%
    group_by(mat_pos) %>%
    add_tally(name = "mat_pos_count") %>%
    left_join(fdrs, by = "mat_pos") %>%
    mutate(SigCount = case_when(
      FDR < 0.01 ~ "Yes",
      TRUE ~ "No"
    ))
}

add_missing <- function(dataset){
  combinations <- crossing(plus = 0:3, minus = 0:2) %>%
    mutate(mat_pos = str_c(plus, minus, sep = ":")) %>%
    anti_join(dataset, by = "mat_pos") %>%
    mutate(n = 0) %>%
    select(mat_pos, n) %>%
    bind_rows(dataset)
}

#step6_test <- step6(step5_test)



```

## Step7: Plot matrix with FDRs and coloring by significance

```{r}
step7_plot <- function(dataset, name, results_dir) {
  print("Executing Step 7")
  temp <- dataset %>%
    ungroup() %>%
    select(-1,-2,-5) %>%
    unique() %>%
    mutate(FDR = round(FDR, digits = 4),
           FDR = case_when(
             .[[1]] == 3 ~ str_c("FDR < ", FDR),
             TRUE ~ str_c("FDR = ", FDR)
           )) %>%
    mutate_all(as.factor) %>%
    ggplot(aes(x = pull(., var = 2), y = fct_rev(pull(., var = 1)), color = SigCount)) +
      geom_text(aes(label = mat_pos_count), size = 8) +
      geom_text(aes(label = FDR), size = 3, nudge_y = -0.3, show.legend = FALSE) +
      theme(panel.grid = element_blank()) +
    labs(x = "no4SU", y = "4SU", title = name)
  
  print(temp)
  ggsave(paste(results_dir, "/SemiQuant_Matrix.png", sep = ""))
  #returning input data for next processing steps
  dataset
}

#step7_test <- step7_plot(step6_test, "0hr, +IFN", "Figures")
```

## Step 8: Find significant proteins from significant peptides

```{r}
step8 <- function(dataset){
  print("Executing Step 8")
  dataset %>%
    ungroup() %>%
    group_by(Proteins) %>%
    mutate(sigTotal = paste(SigCount, collapse = "")) %>%
    mutate(SigCount = case_when(
      str_detect(sigTotal, "No") ~ "",
      TRUE ~ "Yes"
    )) %>%
    select(Proteins, SigCount)
}

#step8_test <- step8(step7_test)
```

## Step 9: Compile final table
 - write to disk
 
```{r}
step9 <- function(full_data, quant, semiquant, results_dir) {
  print("Executing Step 9")
  
  temp <- full_data %>%
    spread(Sample, ave_log_ratio) %>%
    full_join(quant, by = "Proteins") %>%
    full_join(semiquant, by = "Proteins") %>%
    mutate(SigQuant = case_when(
       adj.P.Val < 0.01  & AveExpr > (log2(5)) ~ "Yes",
       TRUE ~ "")
        ) %>%
    select(Proteins, logRatio.Replicate1 = 2,
          logRatio.Replicate2 = 3,
          logRatio.Average = AveExpr,
          pvalue = P.Value,
          pvalue.adjust = adj.P.Val,
          SigQuant = SigQuant,
          SigCount = SigCount,
          -B, -t, -logFC) %>%
    unique() %>%
    mutate(SigQuant = case_when(
      is.na(logRatio.Replicate1) | is.na(logRatio.Replicate2) ~ "",
      TRUE ~ SigQuant
    ))
  
  write_tsv(temp, paste(results_dir, "/Final_Results.tsv", sep = ""))
  
  temp
}

#step9_test <- step9(step3_test, step4_test, step8_test, "Figures")
```
 
## Step 10: Print results summary

```{r}
step10 <- function(data_1, final_data, name){
  print("Executing Step 10")
  
  start_count <- data_1 %>%
    select(Proteins) %>%
    n_distinct()
  
  quant_count <- final_data %>%
    filter(SigQuant == "Yes") %>%
    n_distinct()
  
  semiquant_count <- final_data %>%
    filter(SigCount == "Yes") %>%
    n_distinct()
  
  print(paste("Number of proteins analyzed for ", name, ": ", start_count))
  print(paste("Number of significant proteins by quantitative method: ", quant_count))
  print(paste("Number of significant proteins by semiquantitative method: ", semiquant_count))

}
```

## Run on IAV

```{r}
final_iav <- main(peptides_full, "IAV")
```






## Compare to chikv proteins

```{r, message = FALSE}
# hour0_minus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/0hr, -IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# hour1_minus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/1hr, -IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# hour3_minus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/3hr, -IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# hour0_plus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/0hr, +IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# hour1_plus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/1hr, +IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# hour3_plus <- read_tsv("/Users/saraharcos/Desktop/Ascano Lab/CHIKV_VIRCLASP_Analysis/3hr, +IFN Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# all_chikv <- bind_rows(list("0hr, -IFN" = hour0_minus, 
#                               "1hr, -IFN" = hour1_minus, 
#                               "3hr, -IFN" = hour3_minus, 
#                               "0hr, +IFN" = hour0_plus, 
#                               "1hr, +IFN" = hour1_plus, 
#                               "3hr, +IFN" = hour3_plus), 
#                          .id = "source")
# 
# final_iav <- read_tsv("IAV Results/Final_Results.tsv") %>%
#   filter(SigCount == "Yes" | SigQuant == "Yes")
# 
# chikv_iav <- all_chikv %>%
#   mutate(Virus = "CHIKV") %>%
#   full_join((final_iav %>% mutate(Virus = "IAV")), 
#             by = "Proteins") %>%
#   group_by(Proteins) %>%
#   mutate(Virus = paste(Virus.x, Virus.y, sep = " and "),
#          Virus = str_remove_all(Virus, c(" and NA" = "", "NA and " = ""))) %>%
#   dplyr::select(-Virus.x, -Virus.y) %>%
#   filter(Virus  != "IAV")
# 
# iav_only <- all_chikv %>%
#   mutate(Virus = "CHIKV") %>%
#   full_join((final_iav %>% mutate(Virus = "IAV")), 
#             by = "Proteins") %>%
#   group_by(Proteins) %>%
#   mutate(Virus = paste(Virus.x, Virus.y, sep = " and "),
#          Virus = str_remove_all(Virus, c(" and NA" = "", "NA and " = ""))) %>%
#   dplyr::select(-Virus.x, -Virus.y) %>%
#   filter(Virus  == "IAV") %>%
#   dplyr::select(Proteins)
# 
# ggplot(chikv_iav, aes(x = source, fill = Virus)) +
#   geom_bar(position = "fill")
# 
# library(VennDiagram)
# 
# venn.plot <- venn.diagram(
#   x = list(
#     "IAV" = unique(final_iav$Proteins),
#     "CHIKV" = unique(all_chikv$Proteins)
#   ),
#   filename = NULL,
#   scaled = TRUE,
#   main = "VIR-CLASP Hits"
# 
# )
# 
# pdf(file = "Figures/IAV_CHIKV_Venn_101019.pdf")
#    grid.draw(venn.plot)
#  dev.off()
# 
# 
# overlap <- chikv_iav %>% 
#   filter(Virus == "CHIKV and IAV")
# 
# ggplot(overlap, aes(x = source)) +
#   geom_bar(fill = "steelblue") +
#   #scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25)) +
#   theme_minimal() +
#   theme(text = element_text(size = 16),
#         panel.grid.minor.y = element_blank(),
#         axis.text.x = element_text(angle = 30))
# 
# chikv_iav_all <- all_chikv %>%
#   mutate(Virus = "CHIKV") %>%
#   full_join((final_iav %>% mutate(Virus = "IAV")), 
#             by = "Proteins") %>%
#   group_by(Proteins) %>%
#   mutate(Virus = paste(Virus.x, Virus.y, sep = " and "),
#          Virus = str_remove_all(Virus, c(" and NA" = "", "NA and " = ""))) %>%
#   dplyr::select(-Virus.x, -Virus.y) %>%
#   ungroup()
# 
# genenames = bitr(chikv_iav_all$Proteins, fromType = "UNIPROT", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
# 
# chikv_iav_genenames <- chikv_iav_all %>%
#   full_join(genenames, by = c("Proteins" = "UNIPROT"))
# 
# write_tsv(chikv_iav_genenames, "/Users/saraharcos/Desktop/chikv_iav_101419.txt")

```
```{r}
format_final <- function(dataset){
  names <- bitr(dataset$Proteins, fromType = "UNIPROT", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
  dataset[is.na(dataset)] <- "NA"
  
  dataset %>%
    left_join(names, by = c("Proteins" = "UNIPROT")) %>%
    mutate(`Gene Symbol` = case_when(
      is.na(SYMBOL) ~ Proteins,
      TRUE ~ SYMBOL),
      `Uniprot ID`= Proteins,
      SigCount = case_when(
        SigCount == "NA" ~ " ",
        TRUE ~ SigCount),
      SigQuant = case_when(
        SigQuant == "NA" ~ " ",
        TRUE ~ SigQuant)) %>%
    dplyr::select(`Gene Symbol`, `Uniprot ID`,
           logRatio.Replicate1,
          logRatio.Replicate2,
          logRatio.Average,
          pvalue,
          pvalue.adjust,
          SigQuant,
          SigCount)
}


write.xlsx(format_final(final_iav), file="Final Tables/MaxQuant Final/iav.xlsx", row.names=FALSE)
```










